Probability

### 2016-04-18 prob.tex
I don't know why ShareLaTeX believed I modified a PDF file in it... Hope this time git merge works...

A classical model decides the occurrence or non-occurence of all
events simultaneously which is inconsistent with quantum
mechanics. Indeed, in the quantum world, there are (non-commuting)
events which cannot happen simultaneously. To accommodate this
situation, we completely abandon the sample space $\Omega$ and define
and reason directly about events. Thus a quantum probability space
will consist of just two components: a set of events $\qevents$ and a
probability measure $\phi : \qevents \rightarrow [0,1]$. These
components are defined as follows.

We first assume an ambient Hilbert space $\Hilb$ and define the set of
events $\qevents$ as \emph{projections} on $\Hilb$. Similarly to the
classical case, a probability measure is a function
$\phi : \qevents \rightarrow [0,1]$ satisfying:
\begin{itemize}
\item $\phi(\mathbbm{1}) = 1$, and
\item for all $A \in \qevents$, we have $\phi(A^*A) \geq 0$.
\end{itemize}

As an example, let $P_1, P_2, \ldots, P_k$ be mutually orthogonal
projections on $\Hilb$ with sum $\mathbbm{1}$ and define the event
space $\qevents$ to be the linear span of these operators:
\[
\qevents = \{ \sum_{j=1}^k \lambda_jP_j ~|~ \lambda_i, \ldots, \lambda_k \in \mathbb{C} \}. 
\]
Each state $\ket{\psi}$ of the Hilbert space induces a probability
measure $\phi_\psi : \qevents \rightarrow [0,1]$ defined as follows:
\[
\phi_\psi(A) = \ip{\psi}{A\psi}
\]

%% If we want to follow Swart~\cite{Swart2013} 
%% I am following sec 4 of https://www.math.ru.nl/~maassen/lectures/Trieste.pdf

Concrete example: consider the two qubit Hilbert space with
computational bases $\ket{0}$ and $\ket{1}$ and consider
the following families of projections:
\begin{itemize}
\item Family I: $\proj{0}$, $\proj{1}$
\item Family II: $\proj{+}$, $\proj{-}$
\end{itemize}
and consider the two states \ldots

>>>>

A classical model decides the occurrence or non-occurence of all
events simultaneously which is inconsistent with quantum
mechanics. Indeed, in the quantum world, there are (non-commuting)
events which cannot happen simultaneously. To accommodate this
situation, we completely abandon the sample space $\Omega$ and define
and reason directly about events. Thus a quantum probability space
will consist of just two components: a set of events $\qevents$ and a
probability measure $\phi : \qevents \rightarrow [0,1]$. These
components are defined as follows~\cite{Maassen2010,Swart2013}.

We first assume an ambient Hilbert space $\Hilb$ and define the set of
events $\qevents$ as \emph{projections} on $\Hilb$. Similarly to the
classical case, a probability measure is a function
$\phi : \qevents \rightarrow [0,1]$ satisfying:
\begin{itemize}
\item $\phi(\mathbbm{1}) = 1$, and
\item for all $A \in \qevents$, we have $\phi(A^*A) \geq 0$.
\end{itemize}
\yutsung{If we follow \cite{Maassen2010,Swart2013}, then we also
need 
\begin{itemize}
\item $\phi$ can be extended to a linear functional~$\phi:alg\left(\qevents\right)\rightarrow\mathbb{C}$,
where $alg\left(\qevents\right)$ is the minimal $*$-algebra generated
by $\qevents$.
\end{itemize}
Also, 
\begin{itemize}
\item for all $A\in\qevents$, we have $\phi(A^{*}A)\geq0$. 
\end{itemize}
should be 
\begin{itemize}
\item for all $A\in alg\left(\qevents\right)$, we have $\phi(A^{*}A)\geq0$. 
\end{itemize}
because we have $\phi(A^{*}A)=\phi(A^{2})=\phi(A)$ if $A$ is a projection.}\\
As an example, let $P_1, P_2, \ldots, P_k$ be mutually orthogonal
projections on $\Hilb$ with sum $\mathbbm{1}$ and define the event
space $\qevents$ to be the linear span of these operators:
\[
\qevents=\left\{ \sum_{j=1}^{k}\lambda_{j}P_{j}~\middle|~\lambda_{1},\ldots,\lambda_{k}\in\mathbb{C}\right\} .
\]
\yutsung{$\left\{ \sum_{j=1}^{k}\lambda_{j}P_{j}~\middle|~\lambda_{1},\ldots,\lambda_{k}\in\mathbb{C}\right\} $
is the minimal $*$-algebra generated by $P_{1},P_{2},\ldots,P_{k}$,
but it contains all possible observables $P_{1},P_{2},\ldots,P_{k}$
can generate (and something more) not just projections. For example,
$2\mathbbm{1}\in\left\{ \sum_{j=1}^{k}\lambda_{j}P_{j}~\middle|~\lambda_{1},\ldots,\lambda_{k}\in\mathbb{C}\right\} $.}\\
Each state $\ket{\psi}$ of the Hilbert space induces a probability
measure $\phi_\psi : \qevents \rightarrow [0,1]$ defined as follows:
\[
\phi_\psi(A) = \ip{\psi}{A\psi}
\]
\yutsung{So $\phi_{\psi}$ maps the projections generated by $P_{1},P_{2},\ldots,P_{k}$
to $[0,1]$, and maps $\left\{ \sum_{j=1}^{k}\lambda_{j}P_{j}~\middle|~\lambda_{1},\ldots,\lambda_{k}\in\mathbb{C}\right\} $
to $\mathbb{C}$...}\\

Concrete example: consider the two qubit Hilbert space with computational
bases $\ket{0}$ and $\ket{1}$. First, consider the two states 
\[
\ket{+}=\frac{\ket{0}+\ket{1}}{\sqrt{2}},\ket{-}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}
\]
and consider the following families of projections: 
\begin{itemize}
\item Family I: $\proj{0}$, $\proj{1}$ 
\item Family II: $\proj{+}$, $\proj{-}$
\end{itemize}
In family I, all operators can be expressed as $\left\{ \lambda_{1}\proj{0}+\lambda_{2}\proj{1}~\middle|~\lambda_{1},\lambda_{2}\in\mathbb{C}\right\} $.
In order to identify projectors among them, we need to solve the following
two equations.
\begin{eqnarray*}
\lambda_{1}\proj{0}+\lambda_{2}\proj{1} & = & \left(\lambda_{1}\proj{0}+\lambda_{2}\proj{1}\right)^{*}=\lambda_{1}^{*}\proj{0}+\lambda_{2}^{*}\proj{1}\\
\lambda_{1}\proj{0}+\lambda_{2}\proj{1} & = & \left(\lambda_{1}\proj{0}+\lambda_{2}\proj{1}\right)^{2}=\lambda_{1}^{2}\proj{0}+\lambda_{2}^{2}\proj{1}
\end{eqnarray*}
Therefore, we actually have $\lambda_{1},\lambda_{2}\in\left\{ 0,1\right\} $,
and there are only four projections: $0$, $\proj{0}$, $\proj{1}$,
and $\mathbbm{1}=\proj{0}+\proj{1}$. Then, $\phi_{\ket{+}}$ and
$\phi_{\ket{-}}$ of each projections are 
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\phi_{\ket{+}}(0) & = & 0\\
\phi_{\ket{+}}(\proj{0}) & = & 1/2\\
\phi_{\ket{+}}(\proj{1}) & = & 1/2\\
\phi_{\ket{+}}(\mathbbm{1}) & = & 1
\end{array} & \begin{array}{rcl}
\phi_{\ket{-}}(0) & = & 0\\
\phi_{\ket{-}}(\proj{0}) & = & 1/2\\
\phi_{\ket{-}}(\proj{1}) & = & 1/2\\
\phi_{\ket{-}}(\mathbbm{1}) & = & 1
\end{array}\end{array}
\]
Similarly, $\phi_{\ket{+}}$ and $\phi_{\ket{-}}$ gives two probability
for family II as well.
\[
\begin{array}{c@{\qquad\qquad}c}
\begin{array}{rcl}
\phi_{\ket{+}}(0) & = & 0\\
\phi_{\ket{+}}(\proj{+}) & = & 1\\
\phi_{\ket{+}}(\proj{-}) & = & 0\\
\phi_{\ket{+}}(\mathbbm{1}) & = & 1
\end{array} & \begin{array}{rcl}
\phi_{\ket{-}}(0) & = & 0\\
\phi_{\ket{-}}(\proj{+}) & = & 0\\
\phi_{\ket{-}}(\proj{-}) & = & 1\\
\phi_{\ket{-}}(\mathbbm{1}) & = & 1
\end{array}\end{array}
\]
